\Large \bf Vigor: Toward Carrier-Grade Software Network Functions

### 2016-10-24 doc/paper/paper.tex
paper typesetting: inline code snippets; hyperlinks.

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig}
\usepackage{xcolor}
\usepackage{listings}
\usepackage[]{algorithm2e}
\usepackage{changepage}

\definecolor{mygreen}{rgb}{0.1,0.5,0.1}
\definecolor{myblue}{rgb}{0.1,0.1,0.7}

@@

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\vignat}{{\sc VigNAT}\xspace}
\newcommand{\vignata}{{\sc VigNAT$\alpha$}\xspace}
\newcommand{\vignatb}{{\sc VigNAT$\beta$}\xspace}

@@

In this paper we present work that builds upon these insights and makes three contributions:
\begin{itemize}
    \item {\bf Speculative data plane verification} (SDSE), a technique for verifying stateful software NFs by fusing powerful formal verification with automatic symbolic execution. % It uses formal verification for data structures. It replaces them with simple models that encode some assumptions about the behaviour of the data structures. It then runs SE on the application code and later validates the used models. This approach postpones high-effort/low-return work and discards it when it turns out to be irrelevant.
    \item {\bf The Vigor library of verified components} along with a tool chain that makes SDSE available to those who write software NFs in C/C++. The library consists of data structures (Hash Map, Batcher, Static Array) and a Port Allocator that are common for software NF. % The tool-chain brings together existing SE and formal verification engines to test the proposed technique.
    \item {\bf The \vignat verified NAT box} that offers both high performance and is formally proven to not crash or leak memory. We estimate that it would take an NF developer about 1 week to develop and verify a NAT using Vigor.
\end{itemize}

@@

In order to absorb bursts, the code employs a ring buffer. The interface to the network is: {\em bool receive\_packet(struct packet* dst)}, which non-blockingly reads an inbound packet and stores it at {\em dst}, returning success or failure; {\em bool can\_send\_packet()}, which checks if a new packet can be sent; and {\em void send\_packet(struct packet* src)}, which sends the packet stored at {\em src}. The application follows common practice: it consists of a main event loop, it has no other complex loops, and it uses a ring, a popular data structure in network programming.

@@

One could use an SEE like KLEE~\cite{cadar2008klee} to symbolically execute the code above exhaustively --- once all paths are found to be crash-free, the crash-freedom proof is complete. To perform SE, it is necessary to provide symbolic models for the three network calls. Something simple suffices, such as \code{receive\_packet} generating a new symbolic packet in \code{dst}; \code{can\_send\_packet} forking program state into two states, one for each possible return value; and \code{send\_packet} doing nothing.

@@

For our example, the library provides a \code{ring} with a simple interface: {\em bool ring\_empty()} checks whether there is a packet in the ring, {\em bool ring\_full()} checks whether the ring has run out of space, {\em void ring\_push\_back(struct packet* src)} adds packet {\em src} to the ring, and {\em void ring\_pop\_front(struct packet* dst)} removes a packet from the ring and places it in {\em dst}. We then write a symbolic model for the \code{ring} and provide it to the SEE instead of the actual implementation. The NF now becomes:

@@

%Note, that the \code{ring\_pop} model does not even check whether it is called with an empty ring or not. The example application makes sure to call pop only if the ring is not empty. Therefore the check is not necessary.
%This model clearly overapproximates a ``ring'', e.g. \code{ring\_empty} may return true even right after \code{ring\_push}. This means, it may have caused false positives, but as long as it does not, no need to complicate it.
% For the sake of the example conciseness, we hard coded the invariant ``every packet has port != 9'' and the packet structure into the ring component itself. In a real component this detail should be abstracted and supplied as a parameter.

@@

Of course, execution paths may not always succeed during SE. While this may be indicative of a bug in the NF, it could also be simply because the model is too approximate. For example, in \S\ref{sec:solution-overview} a model of \code{ring\_empty} may be as simple as randomly returning true or false---for some NF code, this behavior could crash it, if it assumes \code{ring\_empty} to return false right after pushing an element. A different model, however, keeps a counter of how many elements are in the ring, and makes \code{ring\_empty} return true only when this counter is zero. The above NF code would not crash with this model; on the downside, the more stateful model makes SE take longer. Therefore, Vigor starts with the least-state models, and tries alternates until it either finds a model with which SE succeeds on all paths or not.

@@

% a number of speculative assumptions that we believe will hold for the real implementations. In the example above some of these assumptions are: ``the \code{ring\_empty} function may return either 0 or 1'', ``if the \code{ring\_pop\_front} returns success the popped packet’s port will never be equal to 9'', and ``\code{ring\_push\_back} never crashes''.

@@

To assess \vignat's latency characteristics, we use the network topology in Figure~\ref{fig:latency-setup}. We use separate ports on the tester for flow generation and latency sampling. We use again DPDK's \code{PktGen} for packet generation. We sample latency with netperf~\cite{jones1996netperf} \code{TCP\_RR}, which measures the request-response delay to \code{netserver} and computes the average over 80 seconds.

>>>>

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig}
\usepackage{xcolor}
\usepackage{listings}
\usepackage[]{algorithm2e}
\usepackage{changepage}
\usepackage{hyperref}
\definecolor{mygreen}{rgb}{0.1,0.5,0.1}
\definecolor{myblue}{rgb}{0.1,0.1,0.7}

\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

@@

\newcommand{\code}[1]{\lstinline{#1}}
\newcommand{\vignat}{{\sc VigNAT}\xspace}
\newcommand{\vignata}{{\sc VigNAT$\alpha$}\xspace}
\newcommand{\vignatb}{{\sc VigNAT$\beta$}\xspace}

@@

In this paper we present work that builds upon these insights and makes three contributions:
\begin{itemize}
    \item {\bf Speculative data plane verification} (SDPV), a technique for verifying stateful software NFs by fusing powerful formal verification with automatic symbolic execution. % It uses formal verification for data structures. It replaces them with simple models that encode some assumptions about the behaviour of the data structures. It then runs SE on the application code and later validates the used models. This approach postpones high-effort/low-return work and discards it when it turns out to be irrelevant.
    \item {\bf The Vigor library of verified components} along with a tool chain
      that makes SDPV available to those who write software NFs in C/C++.
      The library consists of data structures (Hash Map, Batcher, Static Array) and a Port Allocator that are common for software NF. % The tool-chain brings together existing SE and formal verification engines to test the proposed technique.
    \item {\bf The \vignat verified NAT box} that offers both high performance and is formally proven to not crash or leak memory. We estimate that it would take an NF developer about 1 week to develop and verify a NAT using Vigor.
\end{itemize}

@@

In order to absorb bursts, the code employs a ring buffer. The interface to the
network is: \code{bool receive_packet(struct packet* dst)}, which
non-blockingly reads an inbound packet and stores it at \code{dst}, returning success or failure; \code{bool can_send_packet()}, which checks if a new packet can be sent; and \code{void send_packet(struct packet* src)}, which sends the packet stored at \code{src}. The application follows common practice: it consists of a main event loop, it has no other complex loops, and it uses a ring, a popular data structure in network programming.

@@

One could use an SEE like KLEE~\cite{cadar2008klee} to symbolically execute the code above exhaustively --- once all paths are found to be crash-free, the crash-freedom proof is complete. To perform SE, it is necessary to provide symbolic models for the three network calls. Something simple suffices, such as \code{receive_packet} generating a new symbolic packet in \code{dst}; \code{can_send_packet} forking program state into two states, one for each possible return value; and \code{send_packet} doing nothing.

@@

For our example, the library provides a \code{ring} with a simple interface: \code{bool ring_empty()} checks whether there is a packet in the ring, \code{bool ring_full()} checks whether the ring has run out of space, \code{void ring_push_back(struct packet* src)} adds packet \code{src} to the ring, and \code{void ring_pop_front(struct packet* dst)} removes a packet from the ring and places it in \code{dst}. We then write a symbolic model for the \code{ring} and provide it to the SEE instead of the actual implementation. The NF now becomes:

@@

%Note, that the \code{ring_pop} model does not even check whether it is called with an empty ring or not. The example application makes sure to call pop only if the ring is not empty. Therefore the check is not necessary.
%This model clearly overapproximates a ``ring'', e.g. \code{ring_empty} may return true even right after \code{ring_push}. This means, it may have caused false positives, but as long as it does not, no need to complicate it.
% For the sake of the example conciseness, we hard coded the invariant ``every packet has port != 9'' and the packet structure into the ring component itself. In a real component this detail should be abstracted and supplied as a parameter.

@@

Of course, execution paths may not always succeed during SE. While this may be indicative of a bug in the NF, it could also be simply because the model is too approximate. For example, in \S\ref{sec:solution-overview} a model of \code{ring_empty} may be as simple as randomly returning true or false---for some NF code, this behavior could crash it, if it assumes \code{ring_empty} to return false right after pushing an element. A different model, however, keeps a counter of how many elements are in the ring, and makes \code{ring_empty} return true only when this counter is zero. The above NF code would not crash with this model; on the downside, the more stateful model makes SE take longer. Therefore, Vigor starts with the least-state models, and tries alternates until it either finds a model with which SE succeeds on all paths or not.

@@

% a number of speculative assumptions that we believe will hold for the real implementations. In the example above some of these assumptions are: ``the \code{ring_empty} function may return either 0 or 1'', ``if the \code{ring_pop_front} returns success the popped packet’s port will never be equal to 9'', and ``\code{ring_push_back} never crashes''.

@@

To assess \vignat's latency characteristics, we use the network topology in Figure~\ref{fig:latency-setup}. We use separate ports on the tester for flow generation and latency sampling. We use again DPDK's \code{PktGen} for packet generation. We sample latency with netperf~\cite{jones1996netperf} \code{TCP_RR}, which measures the request-response delay to \code{netserver} and computes the average over 80 seconds.
